---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pregnancy247

## Goals
The goal of `pregnancy247` is to process and check the raw sleep diary, Actiwatch, and activPAL data for [The University of Iowa Pregnancy 24/7](https://clinicaltrials.uihealthcare.org/studies/pregnancy-247) research study conducted by [Kara Whitaker](https://clas.uiowa.edu/hhp/people/kara-m-whitaker).

## Installation

You can install the development version of pregnancy247 from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("deboonstra/activpalProcessing")
remotes::install_github("deboonstra/pregnancy247")
```

There is an added installation call to a package called [`activpalProcessing`](https://github.com/deboonstra/activpalProcessing) because this package is no longer available on `CRAN` and an archived copy of this package is used. So, it must be installed prior to installing `pregnancy247`.

## Useage

This package has two unique selections of functions based on the different needs of the Pregnancy 24/7 study team. 

```{r}
library(pregnancy247)
```

There are two main functions in this package, which use the other functions in the package. These functions are `quality_check()` and `process_data()`. The first function checks the quality of the recorded Actiwatch data,

```{r eval = FALSE}
quality_check(subject = "1000-AB", trimester = 1)
```

while the second processes the Actiwatch, sleep diary, and activPAL data.

```{r eval = FALSE}
process_data(subject = "1000-AB", trimester = 1, sleep_source = "./sleep.csv")
```