---
title: "Actiwatch data quality check"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Actiwatch data quality check}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Goal

The goal of `quality_check()` is screen the recorded Actiwatch data and flag outlying values that require further investigation. This document will provide an overview of how to use `quality_check()` appropriately based on the Pregnancy 24/7 study and data structures.

# Load the package
To screen the Actiwatch data to the specifications of the study, use the `library()` command to load the [`pregnancy247`](https://deboonstra.github.io/pregnancy247/articles/pregnancy247.html) package, as shown below.

```{r}
library(pregnancy247)
```

This allows the user to access all of the available functions present in the package, such as `quality_check()`, without a direct reference to the functions, `pregnancy247::quality_check()`.

# Set the working directory
The `quality_check()` function uses the `read_msleep()`, `check_msleep()`, `write_msleep()`, and `write_all_msleep()` functions, and requires that the working directory is set to the directory containing the data of each subject within subject-specific folders based on IDs. It is also assumed the names of the folders in subject-specific directories are

- `Visit 1`,
- `Visit 2`, and
- `Visit 3`.

The requirement of the working directory and file system is based on the needs of the Pregnancy 24/7 study team located at the University of Pittsburgh. Thus, to alter this requirement adjust the dependencies of the file system in the read and write functions. If [RStudio](https://posit.co/products/open-source/rstudio/) is being used, the working directory may be set by the following.

1. Select the *Session* tab in the toolbar
2. Navigate to *Choose Directory ...* listed within *Set Working Directory*
3. Navigate the file explorer to the main data processing folder